---
title: "bellabeat_case_Study"
output: html_notebook 

---

##Business Task

##Discover useful patterns in the data to learn how users are using fitness technology.


##Limitations of the data

##1. No data on gender for a study specifically for women's products. Some women may be held back by biological factors such as menstrual cycles and pregnancy as compared to men's phusical output.

##2.No data on age. Age is known to affect maximum heart rate capable for a human being.

##3. Data is from 2016 so it may be outdated. Popularity in wearable fitness technology soaredin years following with Apple Watch as a big example.


##Product - Fitness Watch

##This wellness watch combines the timeless look of a classic timepiece with smart technology to track user activity, sleep, and stress. The Time watch connects to the Bellabeat app to provide you with insights into your daily wellness.

##Stage 1 - Ask

##1. What observations can be made about user activity? 
##2. How could these observations apply to bellabeat's product?
##3.What should the recommendaed focus points be moving forward?



##Stage 2 - Prepare and Process

---

##Preparing the data

```{r}
activity <- read.csv("Desktop/Fitbit/dailyActivity_merged.csv")
dailysteps <- read.csv("Desktop/Fitbit/dailySteps_merged.csv")
dailycals <- read.csv("Desktop/Fitbit/dailyCalories_merged.csv")
hourlycals <- read.csv("Desktop/Fitbit/hourlyCalories_merged.csv")
weight <- read.csv("Desktop/Fitbit/weightLogInfo_merged.csv")
sleep <- read.csv("Desktop/Fitbit/sleepDay_merged.csv")
```

##Seeing whats in the data.

```{r}
colnames(dailysteps)
```

```{r}
colnames(dailycals)
```

```{r}
colnames(hourlycals)
```

```{r}
colnames(weight)
```

```{r}
colnames(activity)
```

##Structure of largest data set

```{r}
str(activity)
```

```{r}
head(activity)
```

##Checking for Null values in each dataframe I will use.

```{r}
is.null(activity)
```

```{r}
is.null(weight)
```

```{r}
is.null(sleep)
```

##Merging Dataframes

##I will merge the activity data along with the sleep and weight data to find correlation there based in the fact that they are three of the largest factors that lead to a healthy lifestyle.

```{r}
activeweight <- merge(activity, weight, by=c('Id'))
```

```{r}
activesleep <- merge(activity, sleep, by=c('Id'))
```

**Stage 3 - Analysis**

##Analysis

##The longer the distance the more calories burned. This also shows users are using there products because a positive relationship is shown between longer distances and calories burned.

```{r}

ggplot(data = activity) + geom_point(mapping = aes(x = TotalDistance  , y = Calories) , color = 'blue') 
```

##Highly Active users burned the most calories in less minutes. Light Active users burned calories for the most prolonged period of time.

```{r}
ggplot(data = activity) + geom_smooth(mapping = aes(x = VeryActiveMinutes  , y = Calories ) , color = 'blue') +  geom_smooth(mapping = aes(x = FairlyActiveMinutes  , y = Calories ) , color = 'Orange') + geom_smooth(mapping = aes(x = LightlyActiveMinutes  , y = Calories ) , color = 'yellow') + labs(x = "Highly Active (Blue) , Fairly Active (Orange) , Lightly Active (Yellow) ", y = "Calories")
```

##The lighter in weight the more active or perhaps even the more steps taken the lighter one became and the more active as well. The most active individuals weighed between 150 and 200 pounds.

```{r}
ggplot(data = activeweight) + geom_smooth(mapping = aes(x = WeightPounds, y = TotalSteps ))
```

##People who are more active sleep better. We see more sleep in relation to fewer calories burned up until the 3000 when sleep increases likely as a result of increased activity. This hints that being more sedentary leads to more overall sleep but more activity leads to needing more sleep for recovery. The latter likely being a healthier balance at around 8-9 hours of sleep compared to the sedentary 12 hours of sleep.

```{r}
ggplot(data = activesleep) + geom_smooth(mapping = aes (x = Calories, y = TotalTimeInBed))
```

```{r}
ggplot(data = activesleep)+ geom_smooth(mapping = aes(x= SedentaryMinutes, y = TotalMinutesAsleep ))
```

##Summary

##1.There is a positive relationship between activity and calories burned in the user data meaning that the users are engaged with the product.

##2.  More intense activity leads to more calories burned in less time.

##3.There seems to be a dual-relationship between being sedentary and active when it comes to sleep. Sedentary users sleep more likely due to being used to a sedentary life-style. Users who are moderately active sleep a healthy amount but less than those that are sedentary. Active users also see higher durations of sleep likely due to needing recovery.

##4.  Heavier individuals are less active. Individuals between 150-200 pounds are the most active.

**Stage 4 - Act**

##Recommendations

##1.  The watch should have a feature that sets a daily goal for users to meet as well as recommendations of specific goals to meet based on their current physical shape and life style.

##2.  The watch should have sleep recommendations such as a nudge based on user patterns as to when to go to sleep.

##3.  The watch should also recommend types of excercises based on goals. We saw that high intensity resulted in more calories burned in less time. Perhaps incroporating features for High Intensity excercises could be useful for busier individuals.
